\documentclass[a4paper,11pt,twoside]{article}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage[english]{babel}
\usepackage{caption}
\usepackage{cite}
%\usepackage{cleveref}
\usepackage{color}
\usepackage{fancyhdr} % For the page headers
\usepackage{graphicx} % Including images
\usepackage{hyperref} % Including links
\usepackage{etoolbox}
\usepackage{longtable}
\usepackage{sectsty} % Makes changing \chapter, \section, etc. easier

\usepackage{catoptions}

\captionsetup{
  margin=10pt,
  font=small,
  labelfont=bf,
  position=top,
}

\include{includes}

\toggletrue{draft}

\title{Visualization of the projective line geometry for geometric algebra} % FIXME betere titel
\subtitle{Drawing lines in GAViewer}
\author{\href{mailto:dhrdekok@gmail.com}{Patrick de Kok}}
\studentid{5640318}
\ects{18}
\programme{Bacheloropleiding Kunstmatige Intelligentie}
\programmeaddress{University of Amsterdam\\Faculty of Science\\Science Park 904\\1098 XH Amsterdam}
\supervisor{\href{mailto:L.Dorst@uva.nl}{Dr.ir.\ Leo Dorst}}
\supervisoraddress{Intelligent Systems Laboratory Amsterdam\\Faculty of Science\\Science Park 904\\1098 XH Amsterdam}
\date{July 24th, 2012}

%%% Adding own commands, renew old ones
%\renewcommand{\familydefault}{\sfdefault}

\begin{document}
%\frontmatter
\iftoggle{draft}{
\thispagestyle{empty}
\begin{tiny}
\begin{flushright}
Patrick de Kok, 5640318
\end{flushright}
\end{tiny}
  \begin{center}
    \begin{Huge}
      {\color{red} NOTE!}
    \end{Huge}
  \end{center}
  This document contains some notes. Explanation of the notes:
  \begin{itemize}
    \item \colorbox{red}{\bfseries\color{white}Red comments}: directed to the reader in general.  These include questions like how to cite, how something could be rephrased, and questions on spelling.
    \item \colorbox{yellow}{\bfseries Yellow notes}: directed at the supervisor (\makeatletter\@supervisor\makeatother) which are on the subject of the thesis.
    \item \colorbox{niceblue}{\bfseries\color{white}Blue todos}: directed mostly at myself, but might be convenient for the general reader as well.
  \end{itemize}

Now let's start off with a question to the reader:

\comment{If putting a reference at the end of a sentence, how should that be done?}
\begin{enumerate}
  \item This is a reference to some article.~\cite{TheBook}
  \item This is a reference to some article~\cite{TheBook}.
\end{enumerate}

And a small caution:

\TODO{Fix figure and table placement.}

\TODO{Replace $X = \plucker{\V{d}}{\V{m}}$ notation with $x = \plucker{\V{x}}{\V{\bar{x}}} = \V{x}\ez + \edual{\V{\bar{x}}}$}

\TODO{Redefine $X = -\plucker{\V{d}}{\V{m}}$ to $X = \plucker{\V{d}}{\V{m}}$, if possible???}

    \setcounter{page}{-1}
  \newpage
}{}

\maketitle
\begin{abstract}
\input{abstract}
\end{abstract}
\tableofcontents

%\mainmatter
\include{introduction}
\include{background}
\include{research}
\include{results}
\include{conclusion}

\appendix

%\backmatter
\clearpage
\bibliographystyle{plain}
\phantomsection
\addcontentsline{toc}{section}{Bibliography}
\bibliography{citations}

\end{document}
